---
title: "UK Med Losses Scraping"
author: "Mark Biegert"
header-includes:
   - \usepackage{graphicx}
date: "12/2/2021"
output:
  pdf_document: default
  html_document: default
---

# Introduction

This worksheet scrapes UK ship loss data from a web page with OCRed data from an official publicatoini called "British Vessels Lost at Sea, 1935-45", published by HMSO in 1947. The scraping operating was routine, though there were quite a few typos that needed to be fixed.

```{r setup, echo=F, message=F}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(tidyr)
library(dplyr)
library(stringr)
library(lubridate)
library(kableExtra)
library(janitor)
library(rvest)
library(knitr)
library(graphics)
library(kableExtra)
library(ggplot2)
library(hrbrthemes)
library(ggthemes)
library(extrafont)
#font_import()
#loadfonts(device = "win")
#fonttable()
```

```{r global-options, include=FALSE}
   knitr::opts_chunk$set(
        echo = FALSE,
        message = FALSE,
        collapse=FALSE,
        warning = FALSE,
        results='asis'
    )
```

# Background

Drach's comment was about warships. UK losses in the Med are in the many hundreds if you include all the landing craft and auxiliary vessels. I chose to keep only warships starting with the sloop and corvette types. Here are the classes that I kept.

* Aircraft Carrier
* Battleship
* Corvette
* Cruiser				
* Destroyer
* Monitor	
* Sloop		
* Submarine				
		
```{r scrape_losses, echo=F, message=F}
  url = "https://www.naval-history.net/WW2BritishLossesbyArea08.htm"
  general_tmp <- url %>%
  read_html()
  
```

``` {r one_more,echo=F, message=F}
   xpath_ = general_tmp %>%
   html_nodes(xpath="//table/tbody/tr/td/blockquote/p") %>%
   html_text2() %>% 
   str_squish() %>% 
   str_trim()

xpath_ =xpath_[xpath_ != ""][3:462]
xpath_ = xpath_ %>% str_replace("Landing craft assault LCA No.70","January 1 - Landing craft assault LCA No.70")
xpath_= xpath_ %>% str_replace("Landing craft assault LCA No.326","January 1 - Landing craft assault LCA No.326")
xpath_ = xpath_ %>% str_replace("August-September","August 1")
xpath_ = xpath_ %>% str_replace("criser","cruiser")
xpath_ = xpath_ %>% str_replace("BYMS Minesweeper","Minesweeper")
```

```{r form_date, echo=F, message=F, warnings=F}

raw = tibble(src = xpath_)
dsplit = raw %>% separate(src, c("month", "other"), " - ")
ship_type = dsplit %>% 
            mutate(type = word(other,1)) %>% 
            distinct(type)
with_year = dsplit %>% 
            mutate(year = ifelse(is.na(other),month, NA)) %>% 
            fill(year) %>% 
            filter(!is.na(other))

clean_month = with_year %>% 
            separate(month,c("month",NA))

with_date   = clean_month %>% 
              mutate(dt = paste0(month," 1, ",year)) %>% 
              mutate(dt = mdy(dt)) %>% 
              select(date = dt, other)

remove.list = c("Motor","Trawler","Tug","Landing","Yacht","Whaler","Drifter","Human","Water","River","Harbour","Salvage","Boom","Store","Transport","Tanker/oiler","Convoy","Armament","Auxiliary","Small","Armed","Mooring","Schooner","Minelayer ","Minesweeper ")
warships = with_date %>% filter(!word(other,1) %in% remove.list)
parse = warships %>% 
        mutate(name1 = str_replace(other, "([A-Z]{2,})", "|\\1")) %>% 
        mutate(name2 = str_replace(name1," \\(","|\\(")) %>% 
        mutate(name3 = str_replace(name2,"\\),","\\)|")) %>%
        mutate(name4 = str_replace(name3,"P-","|P-")) %>% 
        mutate(name5 = str_replace(name4,"No.","|No.")) %>% 
        mutate(name5 = str_replace(name5, 'Cruiser minelayer','Minelayer')) %>% 
        mutate(name5 = str_to_title(name5)) %>% 
        mutate(name5 = str_trim(name5, side='both'))

parse2 = parse %>% 
        separate(col = name5, into=c('Type','Name',NA),sep='\\|') %>% 
        select(Date=date, Type, Name)  %>% 
        filter(Type!="Submarine Depot Ship ") %>% 
        filter(Type!='Cutter ') %>% 
        filter(Type!="Minelayer ") %>% 
        filter(Type!="Minesweeper ")


```

```{r consolidate, echo=F, message=F}

consolidate = parse2 %>% 
     mutate(Type=str_replace(Type,'Light Cruiser','Cruiser')) %>% 
     mutate(Type=str_replace(Type, 'Escort Destroyer','Destroyer'))

```

# Analysis

## Graphic

I first construct a stacked bar chart to look at what years the ship losses occurred.


```{r graphic, echo=F, message=F, warnings=F}
final = consolidate %>% 
        transmute(Year=year(Date), Type, Name) %>% 
        group_by(Year, Type) 

p1 = ggplot(final, aes(x = factor(Year), fill=factor(Type, levels=c("Monitor ", "Battleship ","Aircraft Carrier ","Sloop ", "Corvette ", "Cruiser ","Mine Warfare ","Submarine ","Destroyer ")))) + 
  geom_bar()+
  guides(fill=guide_legend(title="Ship Type"))+
  
  scale_fill_manual(values = c("Red","Gray","Orange","Black","Brown","Blue","Pink","Green"))+
  ylab('Ships Sunk')+
  xlab('Year')+
  labs(title="UK WW2 Med Losses", subtitle="Breakdown by Year and Ship Type", caption="https://www.naval-history.net/WW2BritishLosses5Support.htm")+
  scale_x_discrete(labels = c('1940','1941','1942','1943','1944'))+
  theme_ipsum()

p1

```

## Tabular Data

The following table shows my summary of the ship losses in tabular form.

```{r tabular, echo=F, message=F}
x = tabyl(final,Type)
y=x %>% arrange(desc(n)) %>% adorn_totals() %>% 
  mutate(percent = scales::percent(percent)) 
y %>%  kable(col.names=c("Type","Number","Percent")) %>%   kable_styling(c("bordered","condensed"),
                latex_options = c("hold_position"), 
                font_size = 14,
                full_width = F)

```

